ms2

### 2017-09-27 ms2/ms2_display.Rmd
ms2: great simplification

- a Spearman correlation of mean claims per beneficiary $\mu$ against mean percent nonsusceptible $R$
- a linear model predicting mean resistance from the fraction $f_>$ of people who are consumers
- a linear model predicting mean resistance from $f_>$ and mean consumption among consumers $\mu_> = \mu/f_>$.

When running regressions, I weight the data points by the number of antibiograms present in that unit. (The weights in a linear regression should be proportional to $1/\sigma^2$, and we generally expect that the standard error of estimators scales like $1/\sqrt{n}$.)

@@

The plot shows the Spearman correlation coefficient $\rho$ between resistance $R$ and mean consumption $\mu$. Two stars after the bug/drug combination implies that the adjusted $p < 0.05$, and one star means the unadjusted $p < 0.05$. Intervals show 95% confidence intervals.

@@

We're interested in how it is that mean consumption drives resistance. Is it at the intensive or extensive margin? We separate that by writing $\mu = f_> \mu_>$.

I run models $R \sim f_>$, that is $R^{(u)} = \beta_0 + \beta_{f_>} f_>^{(u)} + \varepsilon^{(u)}$, where the superscript $u$ indicates the geographical unit (HRR or state). The plots show $\beta_{f_>}$. This model is equivalent to one in which the population resistance $R^{(u)}$ in unit $u$ is the mean of the resistance $r_i^{(u)}$ in the individual people $i$ in unit $u$, where $r_i^{(u)}$ is $\beta_0$ if person $i$ took no drug and $\beta_0 + \beta_{f_>}$ if that person took at least one dose.

The gray line shows $\beta_{f_>} = 1$. Most confidence intervals include that line, i.e., we can explain our data with a transmission-free model where individuals who take drug have 100% resistance and others have some smaller amount. In only one case do we have a place where the data are inconsistent with this simple model. (Note that the $\beta_0$ can mean the resistance among non-consumers or the fraction of non-consumers who are cryptic consumers.)

```{r hrr_linear}
results %>%
  filter(model=='univariate_fnz', unit_type=='hrr', term=='fnz') %>%
  interval_plot() +
  geom_hline(yintercept=1, col='gray') +
  ylab(expression(beta[f['>']])) +
  ggtitle("Linear univariate, HRR")
```

```{r state_linear}
results %>%
  filter(model=='univariate_fnz', unit_type=='state', term=='fnz') %>%
  interval_plot() +
  geom_hline(yintercept=1, color='gray') +
  ylab(expression(beta[f['>']])) +
  ggtitle("Linear univariate, state")
```

Why should the linear models do so much better than the Spearman? I think it's because, in the linear model, I can weight the geographical units that have more antibiograms. That is, we give more weight to the greater amounts of data. The Spearman model treats them all equally.

@@

ineq %>%
  filter(unit_type=='state', drug_group != 'overall') %>%
  ggplot(aes(x=fnz, y=mean, color=drug_group)) +
  stat_smooth(aes(color=NA), color='gray', method='lm', se=F, fullrange=TRUE) +
  geom_point(shape=1) +
  xlab('Fraction nonzero') + ylab('claims per beneficiary') +
  geom_hline(yintercept=0) + geom_vline(xintercept=0) +
  ggtitle('Fraction consuming and mean consumption, states') +
  theme_minimal() +
  scale_color_manual(values=cbbPalette)
```

Because $f_>$ is so collinear with $\mu$, multiple regressions using both as predictors will cause problems (cf. the Supplement). Rather than mean consumption $\mu$, I'll use the mean consumption among consumers $\mu_> = \mu / f_>$ as my second predictor.

```{r fnz_mup}
ineq %>%
  filter(unit_type=='state', drug_group != 'overall') %>%
  ggplot(aes(x=fnz, y=mup, color=drug_group)) +
  geom_smooth(method='lm', se=F) +
  geom_point(shape=1) +
  xlab(expression(f['>'])) +
  ylab(expression(mu['>'])) +
  ggtitle('Mean nonzero consumption vs. fraction consuming, states') +
  theme_minimal() +
  scale_color_manual(values=cbbPalette)
```

@@

```{r fnz_mup_by_drug}
ineq %>%
  filter(unit_type=='hrr') %>%
  #rename(state=unit) %>% left_join(regions, by='state') %>%
  rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
  ggplot(aes(x=fnz, y=mup, color=region)) +
  facet_wrap(~drug_group, scales='free') +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='gray') +
  geom_smooth(method='lm', se=F) +
  geom_point(shape=1) +
  xlab(expression(f['>'])) +
  ylab(expression(mu['>'])) +
  ggtitle('Mean nonzero consumption vs. fraction consuming, states') +
  theme_minimal()
```

Tetracycline and Bactrim are interesting: the South has this long tail where $f_>$ is exceptionally high and $\mu_>$ is low. I imagined a model where the South was like the Midwest, but just with more people taking one does of, say, Bactrim. So it goes from $d$ doses into $c$ consumers to $d + \Delta c$ doses into $c + \Delta c$ consumers. This predicts that $\mu'_> = (\mu_> + \Delta f_>)/(1 + \Delta f_>)$. This prediction is too high: it's not *just* that more people in some parts of the South are getting these drugs; it's also that those getting the drug are getting less of it.

@@

```{r fnz_mean_region}
ineq %>%
  filter(unit_type=='state', drug_group=='overall') %>%
  rename(state=unit) %>%
  filter(state != 'District of Columbia') %>%
  left_join(regions, by='state') %>%
  ggplot(aes(x=fnz, y=mean, color=region)) +
  geom_smooth(aes(color=NULL), method='lm', se=F, color='gray') +
  geom_point() +
  xlab('Fraction nonzero') + ylab('claims per beneficiary') +
  ggtitle('All antibiotics, f> and mean consumption, states') +
  theme_minimal()
```

@@

### Effect along extensive margin $\mu_>$

@@

### Effect beyond $f_>$

We ask if mean consumption $\mu$ has any information beyond fraction nonzero $f_>$ for predicting resistance $R$. We might expect that the simple two-part model (i.e., consuming or not) represented by the regressions $R \sim f_>$ doesn't capture the fact that taking two doses should have a greater effect than taking one, etc.

These plots show the coefficient $\beta_\nu$ from the regression $R = \beta_0 + \beta_{f_>} f_> + \beta_\nu \nu + \varepsilon$.

```{r hrr_multi}
results %>%
  filter(model=='multivariate_fnz_mup', unit_type=='hrr', term=='mup') %>%
  interval_plot() +
  ylab(expression(beta[nu])) +
  ggtitle("Linear multivariate (f> and nu), HRR")
```

At the state level, none of the $\beta_\nu$ coefficients are significantly different from zero (after multiple-hypothesis correction).

```{r state_multi}
results %>%
  filter(model=='multivariate_fnz_mup', unit_type=='state', term=='mup') %>%
  interval_plot() +
  ylab(expression(beta[nu])) +
  ggtitle("Linear multivariate (f> and nu), state")
```

There are a few significant results:

- *E. coli* and Bactrim have a *negative* relationship (higher consumption goes with *less* resistance)
- *S. pneumoniae* and tetracyclines have a negative relationship
- *E. faecium* and $\beta$-lactams have a positive relationship (implying higher mean consumption, given the same fraction of people consuming, goes with more resistance)

Interestingly, these are the same three values that are significant when performing the (ill-advised) regression using $f_>$ and $\mu$. The table shows that bug/drug combinations in which the multiple regressions have significantly non-zero $\beta_{\mu_>}$. The values in the columns show the estimates of $\beta$ in each model. Note the big differences in $\beta_{f_>}$ between the two models. Note that although the regression $R \sim f_> + \mu$ distorts $\beta_{f_>}$ more than the regression $R \sim f_> + \mu_>$, the coefficient still gets changed quite a bit nonetheless. We were hoping for $\mu_>$ to give information in addition to $f_>$, not to just change it completely.

```{r fnz_mu_results}
results %>%
  filter(unit_type=='hrr', model %in% c('univariate_fnz', 'multivariate_fnz_mean', 'multivariate_fnz_mup'), term != '(Intercept)') %>%
  (function(df) {
    filter(df, term != 'fnz', p.adjust(p.value, method='BH') < 0.05) %>%
      select(bug, drug_group) %>%
      distinct() %>%
      inner_join(df, by=c('bug', 'drug_group')) }) %>%
  mutate(model=str_replace(model, '^multivariate_', '')) %>%
  select(bug, drug_group, model, term, estimate) %>%
  spread(term, estimate) %>%
  kable
```

I contrast this situation against *E. coli* and quinolones. In that case, adding terms to the regression doesn't drastically alter $\beta_{f_>}$.

```{r fnz_mu_ecoliquin}
results %>%
  filter(unit_type=='hrr', model %in% c('univariate_fnz', 'multivariate_fnz_mean', 'multivariate_fnz_mup'), term != '(Intercept)') %>%
  filter(bug=='E. coli', drug_group=='quinolone') %>%
  mutate(model=str_replace(model, '^multivariate_', '')) %>%
  select(bug, drug_group, model, term, estimate) %>%
  spread(term, estimate) %>%
  kable
```

For *E. faecium* with $\beta$-lactams, the univariate regression with $f_>$ gives a slightly negative slope but is not significant. The univariate regression with $\mu$ gives a slightly positive slope and is not significant. In the multiple regression, the signs stay the same but the coefficients are very different. Thus, we can't trust that $\mu$ provides data in addition to $f_>$ when it's not clear that either provides the information on its own.

For *S. pneumoniae* and tetracyclines, the regression with $f_>$ and the regression with $\mu$ give negative slopes, but neither are significant. In the multiple regression, the $f_>$ coefficient becomes positive. Again, I'm not sure I can trust this.

For *E. coli* and Bactrim, the univariate regression with $f_>$ is positive and signficant. The regression with $\mu$ is positive and not significant. Then, in the multiple regression, $\mu$ flips signs and becomes highly significant (different from zero), and it's significant in an ANOVA where $f_>$ comes first, but it's not significant in an ANOVA where it ($\mu$) comes first.

@@

From the Spearman and univariate models, Bactrim resistance in *E. coli* is correlated with increased resistance (black line). When breaking that apart by region, however, there are different relationship (colored lines). This makes it seem like there is something strange going on.

It's the L-shape!

```{r ecoli_bactrim_fnz_hrr}
ineq %>%
  filter(unit_type=='hrr') %>%
  rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='tmp_smx') %>%
  filter(percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
  ggplot(aes(x=fnz, y=percent_nonsusceptible, color=region)) +
  geom_abline(intercept=16.86, slope=114.8, color='pink') +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed Bactrim') +
  ylab('Bactrim-resistance E. coli') +
  ggtitle('E. coli & Bactrim across HRRs')
```

These internal trends are a lot weaker when looking at the state level:

```{r ecoli_bactrim_fnz_state}
ineq %>%
  filter(unit_type=='state') %>%
  rename(state=unit) %>%
  left_join(regions, by='state') %>%
  right_join(state_abg, by=c('state', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='tmp_smx') %>%
  ggplot(aes(x=fnz, y=percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed Bactrim') +
  ylab('Bactrim resistance E. coli') +
  ggtitle('E. coli & Bactrim across states')
```

But the same patterns are visible when looking at $\mu_>$ across HRRs:

```{r ecoli_bactrim_mup}
ineq %>%
  filter(unit_type=='hrr') %>%
  rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='tmp_smx') %>%
  filter(percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
  ggplot(aes(x=mean, y=percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('mean Bactrim cons. among Bactrim-eaters') +
  ylab('Bactrim-resistance E. coli') +
  ggtitle('E. coli & Bactrim, mu>, HRRs')
```

@@

```{r pneumo_tetra}
ineq %>%
  filter(unit_type=='hrr') %>%
  rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='S. pneumoniae', drug_group=='tetracycline') %>%
  ggplot(aes(x=fnz, y=percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed tetra') +
  ylab('tetra resistance pneumo') +
  ggtitle('pneumo & tetra across HRRs')

ineq %>%
  filter(unit_type=='state') %>%
  rename(state=unit) %>%
  left_join(regions, by='state') %>%
  right_join(state_abg, by=c('state', 'drug_group')) %>%
  filter(bug=='S. pneumoniae', drug_group=='tetracycline') %>%
  ggplot(aes(x=fnz, y=percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed tetra') +
  ylab('tetra resistance pneumo') +
  ggtitle('pneumo & tetra across states')
```

@@

This negative relationship was significant only for HRRs, not states. Not surprisingly, there is something weird going on with the HRRs: the Southern HRRs are all above the $R \sim f_>$ line, suggesting that the relationship with the mean is just a predictor about region.

```{r ecoli_betalactam}
ineq %>%
  filter(unit_type=='hrr') %>%
  rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='beta_lactam') %>%
  filter(percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
  ggplot(aes(x=fnz, y=percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed beta-lactams') +
  ylab('Beta-lactam resistance in E. coli') +
  ggtitle('E. coli & beta-lactams across HRRs')

ineq %>%
  filter(unit_type=='state') %>%
  rename(state=unit) %>%
  left_join(regions, by='state') %>%
  right_join(state_abg, by=c('state', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='beta_lactam') %>%
  filter(percent_nonsusceptible < 50) %>% # filter one crazy point
  ggplot(aes(x=mean, y=percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed beta-lactams') +
  ylab('Beta-lactam resistance E. coli') +
  ggtitle('E. coli & beta-lactams across states')
```

## Comparing the predictive power of mean consumption and fraction nonzero

How do the regressions $R \sim f_>$ compare to $R \sim \mu$? Here I look at the correlation between the coefficients $\beta_\mu$ and $\beta_{f_>}$ across bug/drug combinations. Not surprisingly, $\beta_\mu$ and $\beta_{f_>}$ are highly correlated: because $\mu$ and $f_>$ are correlated, you can guess the result of the one regression from the other. However, $\beta_{\mu_>}$ is not as well correlated with $\beta_{f_>}$, indicating that there might be some extra information there.

The column titles are `metric1_metric2`, and the value shows the correlation between the univariate models' coefficients for predicting resistance to that drug across bugs.

```{r univariate_result_correlation, eval=FALSE}
results %>%
  filter(str_detect(model, '^uni'), term != '(Intercept)') %>%
  select(unit_type, bug, drug_group, term, estimate) %>%
  spread(term, estimate) %>%
  group_by(unit_type, drug_group) %>%
  filter(n() > 2) %>%
  do(tidy(cor(select(., -unit_type, -bug, -drug_group)))) %>%
  ungroup() %>%
  rename(x=.rownames) %>%
  gather('y', 'value', fnz:nb_mu) %>%
  filter(x > y) %>%
  mutate(key=str_c(x, '_', y)) %>%
  select(-x, -y) %>%
  spread(key, value) %>%
  kable
```

In other words, if you do the regression with $f_>$, you can predict with very high accuracy the coefficient from the univariate regression on $\mu$.

@@

- *E. cloacae* and quinolone
- *S. aureus* and clindamycin

@@

## Collinearity

This shows why it's a bad idea to do a multiple regression with $f_>$ and $\mu$. It confirms that we're mostly getting bad results due to multicollinearity.

The gray lines link the coefficients from a univariate model and a multivariate model (i.e., with $f_>$ and $\mu$) for a certain bug/drug combination.

From a previous plot, we know that $\mu \approx \alpha f_>$, where $\alpha \approx 3/2$. Then if $R \approx \beta_0 + \beta_{f_>}$, then we should guess that $\beta_\mu = \beta_{f_>} / \alpha$.

Furthermore, in doing a multivariate regression, using the same math, we find that we expect
$$
\beta_\mu^{(2)} = \frac{1}{\alpha} \left( \beta_{f_>}^{(1)} - \beta_{f_>}^{(2)} \right),
$$
where the superscripts indicate that the coefficients are from the univariate ("1") and multivariate ("2") regressions. Thus, for $\beta_{f_>}^{(1)} \approx 0$, we expect the multivariate regression coefficients be related by a slope of $-1/\alpha$.

On this plot, the pink lines show $\pm 1/\alpha$, which lines up very nicely with the data (especially considering that I picked $\alpha=3/2$ just by eyeballing that other plot).

```{r uni_multi}
results %>%
  filter(unit_type=='hrr', term != '(Intercept)',
         model %in% c('univariate_fnz', 'univariate_mean', 'multivariate_fnz_mean')) %>%
  mutate(model=str_extract(model, '(uni|multi)'),
         bug_drug=interaction(bug, drug_group)) %>%
  select(bug_drug, model, term, estimate) %>%
  spread(term, estimate) %>%
  ggplot(aes(x=fnz, y=mean, color=model, group=bug_drug)) +
  geom_abline(slope=2/3, color='pink') +
  geom_abline(slope=-2/3, intercept=0, color='pink') +
  geom_line(color='gray') +
  geom_point() +
  xlab(expression(beta[f['>']])) +
  ylab(expression(beta[mu])) +
  ggtitle('Relationship between model coefficients (f>, mean), HRR')
```

But the result is not so pronounced for $f_>$ and $\mu_>$:

```{r uni_multi_mup}
results %>%
  filter(unit_type=='hrr', term != '(Intercept)',
         model %in% c('univariate_fnz', 'univariate_mup', 'multivariate_fnz_mup')) %>%
  mutate(model=str_extract(model, '(uni|multi)'),
        bug_drug=str_c(bug, drug_group)) %>%
  select(bug_drug, model, term, estimate) %>%
  spread(term, estimate) %>%
  ggplot(aes(x=fnz, y=mup, color=model, group=bug_drug, label=bug_drug)) +
  #geom_text(size=2) +
  geom_line(color='gray') +
  geom_point() +
  xlab(expression(beta[f['>']])) +
  ylab(expression(beta[mu['>']])) +
  ggtitle('Relationship between model coefficients (f>, mu>) , HRR')
```

## Using logistic models

We don't expect that, over a large range, the consumption/resistance relationship will be linear. We expect it would be something more like a logistic function. To evaluate the difference between these two approaches, I performed quasibinomial regressions $R \sim f_>$. If $\bar{f}$ is the midpoint of the $f_>$ values (i.e., the weighted mean) and $\hat{y}$ is the predicted resistance from the model, then the slope at the midpoint is $\beta_f \hat{y}(\bar{x}) \left[ 1- \hat{y}(\bar{x}) \right]$. I compare the $p$-values and slopes in the linear and quasibinomial regressions:

```{r linear_quasibin}
results %>%
  filter(unit_type=='hrr', model %in% c('univariate_fnz', 'univariate_fnz_log'), term=='fnz') %>%
  select(bug, drug_group, model, estimate, p.value) %>%
  group_by(model) %>%
  mutate(p.value = p.adjust(p.value, method='BH')) %>%
  ungroup() %>%
  ggplot(aes(x=estimate, y=-log10(p.value), color=model, group=interaction(bug, drug_group))) +
  geom_hline(yintercept = -log10(0.05)) +
  geom_point() +
  geom_line(color='gray')
```

Points above the black line are significant. Note that the $p$-values in the two cases are really similar (i.e., most dumbbells are on the same horizontal line) but they can differ somewhat in their slope. The right-most pink dot is pneumococcus/macrolides, and the next is *P. mirabilis*/Bactrim.

I also tried doing beta regressions, which get similar-sized changes in the slope but really significant $p$-values. Not sure what to make of those.

```{r linear_beta}
results %>%
  filter(unit_type=='hrr', model %in% c('univariate_fnz', 'univariate_fnz_beta'), term=='fnz') %>%
  select(bug, drug_group, model, estimate, p.value) %>%
  group_by(model) %>%
  mutate(p.value = p.adjust(p.value, method='BH')) %>%
  ungroup() %>%
  ggplot(aes(x=estimate, y=-log10(p.value), color=model, group=interaction(bug, drug_group))) +
  geom_hline(yintercept = -log10(0.05)) +
  geom_point() +
  geom_line(color='gray')
```

>>>>

- a Spearman correlation of mean claims per beneficiary $\mu$ against mean percent nonsusceptible $R$
- a linear model predicting mean resistance from the fraction $f$ of people who are consumers
- a linear model predicting resistance from mean consumption among consumers $\nu = \mu/f$
- if both univariate linear models are significant, a linear model predicting resistance from both $f$ and $\nu$

When running regressions, I weight the states/HRRs by the inverse square of the standard error of the unit-level estimates.

@@

The plot shows the Spearman correlation coefficient $\rho$ between resistance $R$ and mean consumption $\mu$. Two stars after the bug/drug combination indicates that the adjusted $p < 0.05$, and one star means the unadjusted $p < 0.05$. Intervals show 95% confidence intervals.

@@

We're interested in how it is that mean consumption drives resistance. Is it at the intensive or extensive margin? We separate that by writing $\mu = f \nu$.

I run models $R \sim f$, that is $R^{(u)} = \beta_0 + \beta_f f^{(u)} + \varepsilon^{(u)}$, where the superscript $u$ indicates the geographical unit (HRR or state). The plots show $\beta_f$. This model is equivalent to one in which the population resistance $R^{(u)}$ in unit $u$ is the mean of the resistance $r_i^{(u)}$ in the individual people $i$ in unit $u$, where $r_i^{(u)}$ is $\beta_0$ if person $i$ took no drug and $\beta_0 + \beta_f$ if that person took at least one dose.

The gray line shows $\beta_f = 1$. Most confidence intervals include that line, i.e., we can explain our data with a transmission-free model where individuals who take drug have 100% resistance and others have some smaller amount. In only one case do we have a place where the data are inconsistent with this simple model. (Note that the $\beta_0$ can mean the resistance among non-consumers or the fraction of non-consumers who are cryptic consumers.)

```{r hrr_linear}
results %>%
  filter(model=='univariate_fnz', unit_type=='hrr', term=='fnz') %>%
  interval_plot() +
  geom_hline(yintercept=0, col='gray') +
  geom_hline(yintercept=1, col='gray') +
  ylab(expression(beta[f])) +
  ggtitle("Linear univariate, HRR")
```

```{r state_linear}
results %>%
  filter(model=='univariate_fnz', unit_type=='state', term=='fnz') %>%
  interval_plot() +
  geom_hline(yintercept=0, color='gray') +
  geom_hline(yintercept=1, color='gray') +
  ylab(expression(beta[f])) +
  ggtitle("Linear univariate, state")
```

@@

ineq %>%
  filter(unit_type=='state', drug_group != 'overall') %>%
  ggplot(aes(x=fnz, y=mean, color=drug_group)) +
  stat_smooth(aes(color=NA), color='gray', method='lm', se=F, fullrange=TRUE) +
  geom_point(shape=1) +
  xlab('fraction nonzero (f)') + ylab('mean consumption (mu)') +
  geom_hline(yintercept=0) + geom_vline(xintercept=0) +
  ggtitle('Fraction consuming and mean consumption, states') +
  theme_minimal() +
  scale_color_manual(values=cbbPalette)
```

Because $f$ is so collinear with $\mu$, multiple regressions using both as predictors will cause problems (cf. the Supplement). Rather than mean consumption $\mu$, I'll use the mean consumption among consumers $\nu = \mu / f$ as my second predictor.

```{r fnz_mup}
ineq %>%
  filter(unit_type=='state', drug_group != 'overall') %>%
  ggplot(aes(x=fnz, y=mup, color=drug_group)) +
  geom_smooth(method='lm', se=F) +
  geom_point(shape=1) +
  xlab(expression(f)) +
  ylab(expression(nu)) +
  ggtitle('Consumption among consumers (nu) vs. fraction consuming (f), states') +
  theme_minimal() +
  scale_color_manual(values=cbbPalette)
```

@@

```{r fnz_mup_by_drug}
ineq %>%
  filter(unit_type=='hrr') %>%
  rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
  ggplot(aes(x=fnz, y=mup, color=region)) +
  facet_wrap(~drug_group, scales='free') +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='gray') +
  geom_smooth(method='lm', se=F) +
  geom_point(shape=1) +
  xlab(expression(f)) +
  ylab(expression(nu)) +
  ggtitle('Mean nonzero consumption vs. fraction consuming, HRR') +
  theme_minimal()
```

Tetracycline and Bactrim are interesting: the South has this long tail where $f$ is exceptionally high and $\nu$ is low. I imagined a model where the South was like the Midwest, but just with more people taking one does of, say, Bactrim. So it goes from $d$ doses into $c$ consumers to $d + \Delta c$ doses into $c + \Delta c$ consumers. This predicts that $\nu' = (\nu + \Delta f)/(1 + \Delta f)$. This prediction is too high: it's not *just* that more people in some parts of the South are getting these drugs; it's also that those getting the drug are getting less of it.

@@

```{r fnz_mean_region}
ineq %>%
  filter(unit_type=='state', drug_group=='overall') %>%
  rename(state=unit) %>%
  filter(state != 'District of Columbia') %>%
  left_join(regions, by='state') %>%
  ggplot(aes(x=fnz, y=mean, color=region)) +
  geom_smooth(aes(color=NULL), method='lm', se=F, color='gray') +
  geom_point() +
  xlab('Fraction nonzero (f)') + ylab('claims per beneficiary (mu)') +
  ggtitle('All antibiotics, f and mean consumption, states') +
  theme_minimal()
```

@@

### Effect along intensive margin $\nu$

@@

### Effect of $f$ beyond $\nu$, or vice versa

We ask if consumption among consumers $\nu$ has any information beyond fraction nonzero $f$ for predicting resistance $R$ (and vice versa). We do regressions $R = \beta_0 + \beta_f f + \beta_\nu \nu + \varepsilon$ and perform $F$-tests on the nested regression models.

```{r double_sig_hrr}
double_sig = results %>%
  filter(model %in% c('univariate_fnz', 'univariate_mup'),
         term != '(Intercept)',
         unit_type == 'hrr') %>%
  group_by(model) %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  ungroup() %>%
  filter(p.value < 0.05) %>%
  group_by(bug, drug_group) %>%
  summarize(n=n()) %>%
  filter(n == 2) %>%
  select(-n)

results %>%
  semi_join(double_sig, by=c('bug', 'drug_group')) %>%
  filter(unit_type == 'hrr',
         model %in% c('multivariate_fnz_mup', 'multivariate_mup_fnz'),
         term != '(Intercept)') %>%
  mutate(second_anova_term=str_extract(model, '(mup|fnz)$')) %>%
  filter(term == second_anova_term) %>%
  select(bug, drug_group, term, anova.p.value) %>%
  spread(term, anova.p.value) %>%
  kable
```

At the HRR level, we see that:

- For CoNS/FQ and *E. cloacae*/FQ, neither $f$ nor $\nu$ gives information beyond the other
- For *E. coli*/FQ and *S. pneumonia*/macrolides, $f$ gives informatio beyond $\nu$
- For *P. mirabilis*/quinolones, $\nu$ gives information beyond $f$
- For *E. coli*/Bactrim, $f$ and $\nu$ each give information beyond the other

And at the state level:

```{r double_sig_state}
double_sig = results %>%
  filter(model %in% c('univariate_fnz', 'univariate_mup'),
         term != '(Intercept)',
         unit_type == 'state') %>%
  group_by(model) %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  ungroup() %>%
  filter(p.value < 0.05) %>%
  group_by(bug, drug_group) %>%
  summarize(n=n()) %>%
  filter(n == 2) %>%
  select(-n)

results %>%
  semi_join(double_sig, by=c('bug', 'drug_group')) %>%
  filter(unit_type == 'state',
         model %in% c('multivariate_fnz_mup', 'multivariate_mup_fnz'),
         term != '(Intercept)') %>%
  mutate(second_anova_term=str_extract(model, '(mup|fnz)$')) %>%
  filter(term == second_anova_term) %>%
  select(bug, drug_group, term, anova.p.value) %>%
  spread(term, anova.p.value) %>%
  kable
```

There are only two bug/drug combinations that are double-significant, and in those cases, neither metric gives information beyond the other.

@@

Increased fraction consuming $f$ is correlated with increased resistance:

```{r ecoli_bactrim_fnz_hrr}
ineq %>%
  filter(unit_type=='hrr') %>%
  rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='tmp_smx') %>%
  filter(percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
  ggplot(aes(x=fnz, y=percent_nonsusceptible, color=region)) +
  #geom_abline(intercept=16.86, slope=114.8, color='pink') +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed Bactrim') +
  ylab('Bactrim-resistance E. coli') +
  ggtitle('E. coli & Bactrim, f, HRRs')
```

But increased consumption among consumers $\nu$ is anticorrelated resistance, both within and across regions:

```{r ecoli_bactrim_mup}
ineq %>%
  filter(unit_type=='hrr') %>%
  rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='tmp_smx') %>%
  filter(percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
  ggplot(aes(x=mup, y=percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('mean Bactrim cons. among Bactrim-eaters') +
  ylab('Bactrim-resistance E. coli') +
  ggtitle('E. coli & Bactrim, nu, HRRs')
```

@@

```{r pneumo_tetra}
ineq %>%
  filter(unit_type=='hrr') %>%
  rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='S. pneumoniae', drug_group=='tetracycline') %>%
  ggplot(aes(x=fnz, y=percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('f') +
  ylab('tetra resistance pneumo') +
  ggtitle('pneumo & tetra, f, HRRs')

ineq %>%
  filter(unit_type=='hrr') %>%
  rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='S. pneumoniae', drug_group=='tetracycline') %>%
  ggplot(aes(x=mup, y=percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('nu') +
  ylab('R') +
  ggtitle('pneumo & tetra, nu, HRRs')
```

@@

```{r ecoli_betalactam}
ineq %>%
  filter(unit_type=='hrr') %>%
  rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='beta_lactam') %>%
  filter(percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
  ggplot(aes(x=fnz, y=percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('f') +
  ylab('Beta-lactam resistance in E. coli') +
  ggtitle('E. coli & beta-lactams, f, HRRs')

ineq %>%
  filter(unit_type=='hrr') %>%
  rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='beta_lactam') %>%
  filter(percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
  ggplot(aes(x=mup, y=percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('nu') +
  ylab('Beta-lactam resistance in E. coli') +
  ggtitle('E. coli & beta-lactams, nu, HRRs')
```

@@

@@
